% This is file JFM2esam.tex
% first release v1.0, 20th October 1996
%       release v1.01, 29th October 1996
%       release v1.1, 25th June 1997
%       release v2.0, 27th July 2004
%   (based on JFMsampl.tex v1.3 for LaTeX2.09)
% Copyright (C) 1996, 1997 Cambridge University Press

\NeedsTeXFormat{LaTeX2e}

\documentclass{jfm}
%\documentclass[referee]{jfm} %for double spaced output for submission


\usepackage{graphicx}
\usepackage{natbib}

\usepackage{color}
\usepackage{soul}
\usepackage{booktabs}
\pdfoptionpdfminorversion 4

% \pdfcompresslevel0

% See if the author has AMS Euler fonts installed: If they have, attempt2
% to use the 'upmath' package to provide upright math.
\ifCUPmtlplainloaded \else
  \checkfont{eurm10}
  \iffontfound
    \IfFileExists{upmath.sty}
      {\typeout{^^JFound AMS Euler Roman fonts on the system,
                   using the 'upmath' package.^^J}%
       \usepackage{upmath}}
      {\typeout{^^JFound AMS Euler Roman fonts on the system, but you
                   dont seem to have the}%
       \typeout{'upmath' package installed. JFM.cls can take advantage
                 of these fonts,^^Jif you use 'upmath' package.^^J}%
       \providecommand\upi{\pi}%
      }
  \else
    \providecommand\upi{\pi}%
  \fi
\fi

% See if the author has AMS symbol fonts installed: If they have, attempt
% to use the 'amssymb' package to provide the AMS symbol characters.

\ifCUPmtlplainloaded \else
  \checkfont{msam10}
  \iffontfound
    \IfFileExists{amssymb.sty}
      {\typeout{^^JFound AMS Symbol fonts on the system, using the
                'amssymb' package.^^J}%
       \usepackage{amssymb}%
       \let\le=\leqslant  \let\leq=\leqslant
       \let\ge=\geqslant  \let\geq=\geqslant
      }{}
  \fi
\fi

% See if the author has the AMS 'amsbsy' package installed: If they have,
% use it to provide better bold math support (with \boldsymbol).

\ifCUPmtlplainloaded \else
  \IfFileExists{amsbsy.sty}
    {\typeout{^^JFound the 'amsbsy' package on the system, using it.^^J}%
     \usepackage{amsbsy}}
    {\providecommand\boldsymbol[1]{\mbox{\boldmath $##1$}}}
\fi

%%% Example macros (some are not used in this sample file) %%%

% For units of measure
\newcommand\dynpercm{\nobreak\mbox{$\;$dyn\,cm$^{-1}$}}
\newcommand\cmpermin{\nobreak\mbox{$\;$cm\,min$^{-1}$}}

% Various bold symbols
\providecommand\bnabla{\boldsymbol{\nabla}}
\providecommand\bcdot{\boldsymbol{\cdot}}
\newcommand\biS{\boldsymbol{S}}
\newcommand\etb{\boldsymbol{\eta}}

% For multiletter symbols
\newcommand\Real{\mbox{Re}} % cf plain TeX's \Re and Reynolds number
\newcommand\Imag{\mbox{Im}} % cf plain TeX's \Im
\newcommand\Rey{\mbox{\textit{Re}}}  % Reynolds number
\newcommand\Pran{\mbox{\textit{Pr}}} % Prandtl number, cf TeX's \Pr product
\newcommand\Pen{\mbox{\textit{Pe}}}  % Peclet number
\newcommand\Ai{\mbox{Ai}}            % Airy function
\newcommand\Bi{\mbox{Bi}}            % Airy function

% For sans serif characters:
% The following macros are setup in JFM.cls for sans-serif fonts in text
% and math.  If you use these macros in your article, the required fonts
% will be substitued when you article is typeset by the typesetter.
%
% \textsfi, \mathsfi   : sans-serif slanted
% \textsfb, \mathsfb   : sans-serif bold
% \textsfbi, \mathsfbi : sans-serif bold slanted (doesnt exist in CM fonts)
%
% For san-serif roman use \textsf and \mathsf as normal.
%
\newcommand\ssC{\mathsf{C}}    % for sans serif C
\newcommand\sfsP{\mathsfi{P}}  % for sans serif sloping P
\newcommand\slsQ{\mathsfbi{Q}} % for sans serif bold-sloping Q

% Hat position
\newcommand\hatp{\skew3\hat{p}}      % p with hat
\newcommand\hatR{\skew3\hat{R}}      % R with hat
\newcommand\hatRR{\skew3\hat{\hatR}} % R with 2 hats
\newcommand\doubletildesigma{\skew2\tilde{\skew2\tilde{\Sigma}}}
%       italic Sigma with double tilde

% array strut to make delimiters come out right size both ends
\newsavebox{\astrutbox}
\sbox{\astrutbox}{\rule[-5pt]{0pt}{20pt}}
\newcommand{\astrut}{\usebox{\astrutbox}}

\newcommand\GaPQ{\ensuremath{G_a(P,Q)}}
\newcommand\GsPQ{\ensuremath{G_s(P,Q)}}
\newcommand\p{\ensuremath{\partial}}
\newcommand\tti{\ensuremath{\rightarrow\infty}}
\newcommand\kgd{\ensuremath{k\gamma d}}
\newcommand\shalf{\ensuremath{{\scriptstyle\frac{1}{2}}}}
\newcommand\sh{\ensuremath{^{\shalf}}}
\newcommand\smh{\ensuremath{^{-\shalf}}}
\newcommand\squart{\ensuremath{{\textstyle\frac{1}{4}}}}
\newcommand\thalf{\ensuremath{{\textstyle\frac{1}{2}}}}
\newcommand\Gat{\ensuremath{\widetilde{G_a}}}
\newcommand\ttz{\ensuremath{\rightarrow 0}}
\newcommand\ndq{\ensuremath{\frac{\mbox{$\partial$}}{\mbox{$\partial$} n_q}}}
\newcommand\sumjm{\ensuremath{\sum_{j=1}^{M}}}
\newcommand\pvi{\ensuremath{\int_0^{\infty}%
  \mskip \ifCUPmtlplainloaded -30mu\else -33mu\fi -\quad}}

\newcommand\etal{\mbox{\textit{et al.}}}
\newcommand\etc{etc.\ }
\newcommand\eg{e.g.\ }


\newcommand{\uu}{\textbf{u}}
\newcommand{\xx}{\textbf{x}}
\newcommand{\kk}{\textbf{k}}
\newcommand{\JJ}{\textbf{J}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\ff}{\textbf{f}}
\newcommand{\FF}{\textbf{F}}
\newcommand{\baa}{\textbf{a}}
\newcommand{\bb}{\textbf{b}}
\newcommand{\NN}{\textbf{N}}


\newcommand{\eps}{\varepsilon}


\newcommand{\ErtelPV}{\mathcal{Q}}

\newcommand{\fdiss}{f_{\mbox{\scriptsize diss}}}

\newcommand{\D}{\mbox{D}}

\newcommand{\eez}{\boldsymbol{e_z}}
\newcommand{\scalarprod}[2]{\big( #1 \, , \ #2 \big)_{\kk}}

\newcommand{\mean}[1]{\langle #1 \rangle}

\newcommand{\meane}[1]{\langle #1 \rangle}
% \newcommand{\meane}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}e}}}

\newcommand{\meanx}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}\mbox{\scriptsize$\xx$}}}}

\newcommand{\meant}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}\theta}}}

\newcommand{\means}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}s}}}

\newcommand{\shocks}{ {\mbox{\tiny shocks}} }

\newcommand{\kmax}{k_{\mbox{\scriptsize max}}}
\newcommand{\kdiss}{k_{\mbox{\scriptsize diss}}}


\newcommand{\mA}{\mathcal{A}}


\newlength{\halfwidth}
\setlength{\halfwidth}{2.6in}

\newcommand{\PA}[1]{{\color{green}#1}}

\newcommand{\Add}[1]{{\color{blue}#1}}
% \newcommand{\Add}[1]{#1}
\newcommand{\Remove}[1]{{\color{red}\st{#1}}}
% \newcommand{\Remove}[1]{}

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\title[Shallow water wave turbulence]%
{ Shallow water wave turbulence}

\author[P. Augier, A.V. Mohanan and E. Lindborg]%
{Pierre Augier$^{1}$,   A. V. Mohanan$^{2}$ 
and Erik Lindborg$^2$ \ns }

\affiliation{$^2$ KTH Mechanics,
SE-100 44 Stockholm, Sweden\\[\affilskip]
$^1$ LEGI, BP53,
38041 Grenoble Cedex, France}

\pubyear{2013}
\volume{???}
\pagerange{???--???}
% Do not enter received and revised dates. These will be entered by
% the editorial office.
\date{?; revised ?; accepted ?. - To be entered by editorial office}
%\setcounter{page}{1}
\begin{document}

\maketitle


% \baselineskip 8mm
\begin{abstract}
The dynamics of irrotational shallow water wave turbulence forced in large scales
and dissipated  at small scales is investigated.
First, we derive the shallow water analogue of the `four-fifths law' of three-dimensional turbulence for a third order structure function involving velocity and displacement increments.
Using this relation and assuming that the flow is dominated by shocks we develop a simple model predicting that the shock amplitude scales as $ (\epsilon d)^{1/3} $, where $ \eps $ is the mean dissipation rate and $ d $ the mean distance between the shocks, and that the
$ p $:th order displacement and velocity structure functions scale as
$ (\eps d)^{p/3} r/d $, where $ r $ is the separation. 
Then we carry out a series of forced simulations with resolutions up to $7680^2$, varying the Froude number,  $F_{f} = \eps^{1/3}  / ck_f^{1/3} $, where $ k_f $  is the forcing wave number and $c$ is the wave speed.
In all simulations a stationary state is reached  in which there is a constant spectral energy flux and equipartition between kinetic and potential energy in the constant flux range. The third order structure function relation is satisfied with a high degree of accuracy. 
Mean energy is found to scale as $ E \sim \sqrt{\epsilon c/k_f} $,  and is also dependent on resolution, indicating that shallow water wave turbulence does not fit into the paradigm of a Richardson-Kolmogorov cascade.  
In all simulations shocks develop, displayed as long thin bands of negative divergence in flow visualizations. The mean distance between the shocks is found to scale  as $ d \sim F_f^{1/2}/k_f $. Structure functions of second and higher order are found to scale in good agreement with the model. We conclude that in the weak limit, $ F_f \rightarrow 0 $, shocks will become denser and weaker and finally disappear for a finite Reynolds number. On the other hand, for a given $ F_{f} $, no matter how small, shocks will prevail if the Reynolds number is sufficiently large. 
 

\end{abstract}

% \begin{keywords}
% \end{keywords}

\section{Introduction}

The shallow water (SW) equations  have been widely used to study basic
mechanisms occurring in geophysical flows \cite[see for
example][]{VallisLIVRE2006}.
%
The equations are based on the hydrostatic approximation which is
extremely well satisfied for a very wide range of scales in the oceans
and in the atmosphere.  They also rely on a stronger hypothesis
regarding the density stratification.  The fluid is assumed to be
structured in a limited number of thin homogeneous layers, which leads
to equations involving only two-dimensional operators. The two-layer
SW equations capture the baroclinic instability which is of
primary importance for the dynamics of the oceans and of the
atmosphere \cite[]{VallisLIVRE2006, Wirth2013}.
%
The single layer SW equations are very similar to
the two-dimensional compressible Navier-Stokes equations.  The
equations do not capture baroclinic instability but they constitute 
one of the simplest hydrodynamic models with coexisting eigenmodes of
the linearized operator with zero linear frequency and with finite
non-zero linear frequency, $\omega  \pm\sqrt{f^2 + c^2
k^2}$, where $f$ is the Coriolis parameter, $c$ the gravity wave speed and $k
= |\kk|$ the wave number.
%

For geophysical applications, it is convenient to consider the limit
for which the linear wave frequency  is much larger than the
characteristic nonlinear frequency $ U k$, where $U$ is a
characteristic velocity.  In this limit, one can obtain a quasi-geostrophic
single layer SW model, i.e. derive self-consistent evolution
equations for a zero frequency variable, the Charney potential
vorticity, which is totally decoupled from the fast waves.
%
For finite rotation and wave speed, there are slow balanced motions,
which are associated with relatively weak non-zero linear frequency
modes and there is only a weak
interaction between waves and the balanced motions.
%
Due to their relative simplicity, the SW equations have
been used to study issues like the projection on the balanced manifold
\cite[]{Lorenz1980, MohebalhojehDritschel2000} and the production of
waves by balanced flows \cite[]{FargeSadourny1989,
LahayeZeitlin2012,Vanneste2013}. 
In this study, we consider the
same equations in order to address a related but different issue.

Balanced flows have a dynamics which is very similar to
two-dimensional turbulence, with an upscale energy cascade. %
Small scale energy dissipation vanishes in the limit of zero
viscosity.
%
In contrast to this, statistical mechanics indicates that wave energy
should be transferred towards small scales \cite[]{Warn1986}.
Numerical simulations have shown that long waves lose their energy by
transferring it to shorter waves \cite[]{Sadourny1975,
FargeSadourny1989, YuanHamilton1994}.  The mechanism of a wave cascade
has been proposed to explain the attraction towards the balanced flows
\cite[]{Sadourny1975}.  \cite{YuanHamilton1994} showed that
statistically stationary shallow-water flows can be obtained forcing
quasi-geostrophic modes at large scales and with dissipation only at
small scales.
%
They propose that the $k^{-5/3}$ spectra that are observed at the
atmospheric meso-scales  \cite[]{NastromGage1985,LiLindborg2018}
and reproduced by some atmospheric models based on the primitive
equations \cite[]{KoshykHamilton2001, Skamarock2004,
Hamilton_etal2008} could be reproduced by the one-layer
shallow-water model by a downscale energy cascade of waves.



However, few studies have investigated the dynamics of the downscale
energy cascade of waves in SW models.
%
Most numerical studies have investigated decaying flows and have not
focused on the cascade \cite[]{FargeSadourny1989,
LarichevMcWilliams1991, SpallMcWilliams1992,
PolvaniMcWilliamsSpallFord1994, LahayeZeitlin2012}.
%
Moreover, there are strong discrepancies between different numerical
results and between these results and available theoretical results.
%
\cite{FargeSadourny1989} reported spectra scaling as $k^{-4}$
whereas those shown by \cite{LahayeZeitlin2012} scale as $k^{-6}$.

Some predictions are based on the weak wave turbulence formalism
\cite[]{ZakharovLvovFalkovich1992, Nazarenko2011}.  In the case of
inertia-gravity waves, \cite{FalkovichMedvedev1992} showed that the
exact solutions of the kinetic equation corresponding to a constant
downscale energy flux is associated with a spectrum scaling as $E(k)
\sim  k^{-8/3}$, while \cite{ZakharovSagdeev1970} 
used weak turbulence formalism to derive a spectrum of the form $ E(k) \sim  k^{-3/2} $ for acoustic turbulence. This prediction is of particular interest, since the aqoustic equations and the SW equations are identical when a first order expansion is made in
density or displacement pertubations. It may be questioned, however,  that weak turbulence formalism is applicable to acoustic and SW wave turbulence, no matter how weak the nonlinearities are supposed to be. As pointed out by \cite{KadomtsevPet1970} and 
\cite{FalkovichMeyer1996}, when dispersion is small enough, waves moving in the same direction have
the same velocity and interact strongly, which leads to the
formation of shocks.  In this case, the weak wave turbulence
formalism is unable to give valuable results.  Such sharp and long
structures have been observed in many simulations
\cite[]{FargeSadourny1989, PolvaniMcWilliamsSpallFord1994,
LahayeZeitlin2012} but their effects are not well understood.
%
As shown by \cite{Burgers1948} the energy spectrum of a flow dominated by shocks will scale as $ k^{-2} $, a 
 prediction which has been confirmed by a great number of simulations of Burgers equation \cite[for a review, see ] []{Frisch2001}.
 The shocks also lead to very strong intermittency, as demonstrated by
\cite{BouchaudMezardParisi1995} and \cite{WeinanKhaninMazelSinai1997}
for Burgers turbulence.
%
\cite{FalkovichMeyer1996} simulated acoustic turbulence dominated by shocks 
and reported spectra scaling as $ k^{-2} $.
\cite{YuanHamilton1994} carried out
forced dissipative simulations of the SW equations showing similar spectra. 
However, the resolution in both these studies was quite coarse, ($386^2$
and $101^2$ grid points, respectively) so that the scaling range was very
narrow.  Moreover, \cite{FalkovichMeyer1996} used a very particular
type of strongly anisotropic forcing with waves in only one
direction while \cite{YuanHamilton1994} forced in geostrophic modes
which severely complicates the interpretation.

The dynamics of waves in complex SW flows is not well
understood.  Even wave dynamics in absence of a balanced flow is not well
understood. A statistically stationary wave cascade has never been
simulated at high resolutions.
%
Is it possible to obtain a statistically stationary state in the very
fast wave regime, and how does the mean energy in this hypothetical
regime depends on the forcing and the wave speed?  
% What is the slope of 
How do energy spectra scale with the 
input parameters? From a theoretical point of view, the most interesting issue is, perhaps,  the nature of the dynamics in the limit of weak nonlinearities. Will shocks prevail in this limit?
%
Such questions are still open and straightforward simulations may
provide the answers.
In this paper, we investigate the dynamics of single layer SW wave turbulence 
forced  at large scales by means of 
simulations with up the  $7680^2$ grid points.
%
Before presenting the results of the simulations we develop some theory.
%


\section{Theory}
\label{section_basic_theory}
\input{section_basic_theory}


\label{section_advanced_theory}
\input{section_advanced_theory}

\section{Simulations}
\label{section_methods}
\input{section_methods}


%\label{section_wave_cascade_f=0}
%\input{section_wave_cascade_f=0}

%\section{Effects of global rotation}
%\label{section_rotation}
%\input{section_rotation}



\section{Conclusions}

We have carried out a theoretical and numerical investigation of SW wave turbulence. First, we derived the SW analogue (\ref{eq_Kolmo}) of the the `four-fifths' law of Kolmogorov turbulence. Using this relation and straightforward statistical and geometrical arguments we developed a simple shock model predicting that the shock amplitude scales as $ (\epsilon d)^{1/3} $, where $ d $ is the mean distance between the shocks, and that the $ p $:th order structure function above a certain minimum will scale as $ (\epsilon d)^{p/3} r/d $. Then, we carried out a series of twenty forced dissipative simulations, varying the Froude number and the resolution. In all simulations, a statistically stationary state were reached and in all simulations we observed shocks. The flow variables in each Fourier mode was found to evolve as a linear wave, with equipartition between APE and KE over a period. The third order structure function relation was fulfilled with a high degree of accuracy. From the simulations we made two interesting observations that fall outside the predictive range of the model. The first observation  was  that mean energy in the stationary state approximately scales as (\ref{MeanEnergy}), suggesting that the normalised dissipation (\ref{Dissipation}) will go to zero both in the limit of small viscosity and large  wave speed. This is different from three dimensional Kolmogorov turbulence as well as strongly stratified turbulence, where, in both cases, dissipation is finite in the limit of small viscosity, and in the latter case also is independent of the degree of stratification. This observation suggests that SW wave turbulence does not fit into the paradigm of a local Richardson-Kolmogorov cascade. The second observation we made was that the mean distance between the shocks scales as $ c^{-1/2} $, if the other parameters are held constant. As we increased $ c $, the shocks were thus becoming denser. Using this observation we tested the shock model by plotting the normalised and compensated structure functions of different orders. Generally we found that there was a quite good agreement between the simulation results and the model predictions, becoming better with decreasing  wave speed and increasing resolution. 

Combining the model and the observations we can get a quite good picture of the dynamics in the limit of weak nonlinearities, $ F_{f} \rightarrow  0 $.  
The picture may become more vivid if we say `wave breaking' were we previously have talked about shock formation, and also say
`wave lengths' were we previously have talked about length scales. We identify the longest wave length, $ \lambda_l $,  with the forcing scale $  L_f $ and the breaking wave length, $ \lambda_b $, with the mean distance, $ d $, between the shocks. There are two dynamically different  regimes to consider, the short wave  regime, $ [\delta x, \; \lambda_b] $, and the long wave regime, $ [\lambda_b, \, \lambda_l] $. We first consider the latter, whose width scales  $ \lambda_{l} /\lambda_b \sim F_f^{-1/2} $. Identifying the amplitude of the breaking waves with the shock amplitude, $ (\eps d)^{1/3} $, we find that the ratio between the amplitude of the longest waves and the breaking waves scales as $ E^{1/2}/(\eps d)^{1/3} \sim F_{f}^{-5/12} n^{\alpha/2} $, where we may substitute  $ n $ with a Reynolds number to some power.  Thus, for small $ F_f $ and large Reynolds number the longest waves are not directly influenced by wave breaking. The Reynolds number dependence of the amplitude ratio makes it unlikely that a general similarity theory can be easily formulated for this range. Therefore, predictions from weak turbulence theory \cite[for example][]{ZakharovSagdeev1970}
are not likely to hold for this range, although it is not directly influenced by wave breaking. The exact values of the power law exponents in the argument, are, of course, not important. As for the short wave regime, the width scales as $  \lambda_b /\delta x \sim F_f^{1/2} $ for fixed $ \delta x $ or fixed Reynolds number. As $ F_{f} \rightarrow 0 $ the range will eventually become so narrow so that breaking will disappear. On the other hand, it may be argued that wave breaking will take place for any small $ F_f $, provided that the Reynolds number is sufficiently large. Apparently, a necessary condition for wave breaking is that
\begin{equation} \label{NC}
\frac{\lambda_b}{\delta x} \sim \frac{\lambda_l F_f^{1/2}}{\delta x} \gg 1 \, .
\end{equation}
It may be interesting to note that the single point limit of the flatness factors shown if figure (\ref{fig_flatness}) scales as the ratio $ \lambda_b/\delta x $ and that this limit is equal to the flatness factor of velocity derivatives. So the condition (\ref{NC}) means that the flatness factor of velocity derivatives must be much larger than unity. Assuming that this is not only a necessary but also sufficient condition, we would able to formulate under what condition wave breaking will take place in terms of a Froude number and a Reynolds number if we knew how the shock width $ \delta x $ scales with these parameters.
In setting up the simulations we assumed that it is sufficient to resolve the Kolmogorov scale, $ \eta = (\nu^{3} / \epsilon)^{1/4} $, or the corresponding scale, $ \eta_8 = (\nu_8^{3}/\epsilon) ^{1/22} $, when hyperviscosity is used. The results from the simulations seem to support this assumption, suggesting that $ \delta x $ can be identified as the Kolmogorov scale.  Assuming that this is correct and defining the Reynolds number as $ Re \equiv \lambda_l^{4/3}\epsilon ^{1/3} /\nu $, we can write the condition for wave breaking as
\begin{equation} \label{Condition} 
Re^{3/4} Fr^{1/2} \gg 1 \, .
\end{equation} 
However, a little bit of reflection may lead us to doubt that the shock width is equal to the Kolmogorov scale. The reason for this is that dissipation only occurs at the shocks.
Using Navier-Stokes viscosity and assuming that the shock width is the characteristic dissipation length scale, the mean energy dissipation rate can be estimated as
\begin{equation}
\eps \sim \nu \frac{\mid \Delta u \mid ^2}{\delta x ^2} \frac{L_s \delta x}{\cal{A}} \sim \nu \frac{ \;\;\;  (\epsilon \lambda_b)^{2/3} }{\delta x \lambda_b} \, ,
\end{equation} 
where we have used the estimate (\ref{Strength}) for the shock amplitude. Consequently we obtain
\begin{equation} \label{ShockWidth}
\delta x \sim \frac{\eta^{4/3}}{\lambda_b^{1/3}} \sim \frac{\eta^{4/3}}{\lambda_l^{1/3}} F_f^{-1/6} \, ,
\end{equation} 
where we have used the observed scaling law (\ref{MeanDistance}) for $ \lambda_b = d $. An analogous result is obtained if we instead use hyperviscosity.  The estimate (\ref{ShockWidth}) suggests that $ \delta x $ should become somewhat larger  with increasing $ c $, but very slowly, $ \delta x \sim c^{1/6} $.  It is difficult to judge from the simulations  if this is actually the case. In figure (\ref{fig_spectra_c40}) we see that the maximum of the bottleneck in the high wave number end of the spectra appears at a somewhat smaller wave number in the run with largest $ c $ as compared to the two other runs. We leave it as an open question whether this is a signature of a slowly increasing shock width with increasing $ c $. It is interesting to note that if (\ref{ShockWidth}) is used as the estimate of $ \delta x $ in (\ref{NC}) instead of the Kolmogorov scale, we actually obtain the same condition (\ref{Condition}).

The question that originally motivated this study was if SW wave turbulence can be used as a model for energy dissipation in the the atmosphere and the oceans. Undoubtedly, the answer to this question is negative and the reason for this is not primarily the appearance of shocks and the associated $ k^{-2} $-spectrum which is not consistent with the observed $ k^{-5/3} $-spectrum. The main reason is that 
SW wave turbulence is  a far too inefficient agent of dissipation. Since the Reynolds number of geophysical flows is extremely large and the stratification is strong, the  relation (\ref{Dissipation}) implies that dissipation would be absent in the interior of the atmosphere and the oceans with SW turbulence as the responsible agent. This not what we observe.

Although the answer to the question that originally motivated this study is negative, it may lead us or some other investigators to pursue another interesting research direction. As pointed out in the introduction, the acoustic equations are to the lowest order expansion in density pertubations identical to the SW equations. This study may therefore give important clues to the long standing problem of acoustic turbulence, in two and three dimensions. A similar shock model as we developed in two dimensions may easily be formulated in three dimensions, where the third order structure function law will be  similar to (\ref{eq_Kolmo}), apart from the constant. In three dimensions we find
\begin{equation}
\meane{ |\delta \uu|^2 \delta J_L } 
+ c^2\meane{ (\delta h)^2 \delta u_L } = - \frac{8}{3} \varepsilon r. \label{eq_Kolmo3}
\end{equation}
 In the case of acoustic turbulence $ \delta h $ should  be interpreted as a density increment rather than a displacement increment.  If shocks in  three dimensions are assumed to form as  smooth surfaces rather than smooth lines, then the mean distance, $ d $, between the shocks will appear in the model as $ d \sim {\cal{V}}/A_s $, where $ A_{s} $ is the total  shock area and $ \cal{V} $ is the volume of the domain. The shock amplitude will thus scale as $ (\epsilon d)^{1/3} $ and the structure functions will scale as $ (\epsilon d)^{p/3} r/d $ also in three dimensions, provided, of course, that shocks will form in a similar way as in two dimensions. If they do, the most interesting questions are if the mean energy and the mean distance between the shocks scale in a similar way in three dimensions as in two dimensions. Simulations or experiments of acoustic turbulence may provide answers to these questions -- answers that will lead to further clues to a theoretical understanding of acoustic turbulence. 
 We hope that the present study will stimulate research in this direction.  
 


%\appendix

%\section{Kinetic energy dissipation and forcing rates}
%\label{app_comp}
%\input{appendix_dissip_forcing}



\bibliographystyle{jfm}
\bibliography{biblio}

\end{document}

